---
output: pdf_document
---
# Analysis 11 - Comparing clusters of Wildtype large

Based off of `lcmSOM_analysis10_122916_redo`, with newly mapped genes.

## Purpose 

Comparing the clusters for two purposes

1. Quality of Analysis - To understand how consistent different clusteres are, therefore how conservative the results are.

2. Decrease the amount of interesting genes in clusters by comparing WT SOM with *tf2* superSOM. 

## Analysis Set-up

Required Libraries

```{r}
library(ggplot2)
library(reshape)
library(kohonen)
library(goseq)
library(GO.db)
library(knitr)
library(somtools)

## all the somtools functions
ls("package:somtools")

## all the functions in script form
#source("./clusterFunctions.R")
```

Upload that dataset:

```{r}
genes25 <- read.csv("../data/output/analysis4.top25_19Oct2017.csv")
colnames(genes25)

## Only WT
genes25 <- genes25[,c(2,9:14)]
```

```{r}
scale_data <- as.matrix(t(scale(t(genes25[c(2:7)])))) 
pca <- prcomp(scale_data, scale = TRUE) 
summary(pca) 
pca.scores <- data.frame(pca$x)
data.val <- cbind(genes25, scale_data, pca.scores) 

PCA12 <- ggplot(data.val, aes(PC1, PC2)) +
  geom_point(alpha = .3)

PCA34 <- ggplot(data.val, aes(PC3, PC4)) +
  geom_point(alpha = .3)

plot_grid(PCA12, PCA34, labels = c("", ""), ncol = 2)
```

## SOM 

```{r}
# subset only the scaled gene expression values
# colnames(scale_data) <- NULL

set.seed(6)
som <- som(scale_data, somgrid(6,6)) # This is where you change the size of the map
summary(som)

lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
lcmPalette <- function(n, alpha =1){
  lcmPaletteColors[1:n]
}
  
plot(som, type = "codes", palette.name = lcmPalette)

## Bind information
data.val2 <- cbind(data.val,som$unit.classif,som$distances) 
names(data.val2)
##rename 
names(data.val2)[20:21] <- c("unitClassIfA", "distanceA")
```

## Visualization

```{r}
# Palette
lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
```

## Summary

Just want to visualize the size of the different clusters.
```{r}

str(data.val2)
sum_clusters <- as.data.frame(summary(as.factor(data.val2$unitClassIfA)))

colnames(sum_clusters)[1] <- "total"
sum_clusters$clusters <- row.names(sum_clusters)

ggplot(sum_clusters, aes(reorder(clusters, -total), total)) + 
    geom_histogram(stat = "identity")

mean(sum_clusters$total)
```

## Cluster 1
```{r}
clusterVis(1)
clusterVis_region(1)
clusterVis_color(1)
clusterVis_PCA(1) 
y <- genesInClust(1, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 2
```{r}
clusterVis(2)
clusterVis_region(2)
clusterVis_color(2, lcmPaletteColors)
clusterVis_PCA(2)
clusterGO(2)
y <- genesInClust(2, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```


## Cluster 3

```{r}
clusterVis(3)
clusterVis_region(3)
clusterVis_color(3, lcmPaletteColors)
clusterVis_PCA(3)
clusterGO(3)
y <- genesInClust(3, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 4

```{r}
clusterVis(4)
clusterVis_region(4)
clusterVis_color(4, lcmPaletteColors)
clusterVis_PCA(4)
clusterGO(4)
y <- genesInClust(4, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

##Cluster 5

```{r}
clusterVis(5)
clusterVis_region(5)
clusterVis_color(5, lcmPaletteColors)
clusterVis_PCA(5)
clusterGO(5)
y <- genesInClust(5, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

##Cluster 6

```{r}
clusterVis(6)
clusterVis_region(6)
clusterVis_color(6, lcmPaletteColors)
clusterVis_PCA(6)
clusterGO(6)
y <- genesInClust(6, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 7
cluster 

8 - up-regulated in margin
9 - down in expression along longitudinal axis - cell cycle 

```{r}
clusterVis(8)
clusterVis_region(8)
clusterVis_color(8, lcmPaletteColors)
clusterVis_PCA(7) 
clusterGO(7)
y <- genesInClust(9, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 7
cluster 

8 - up-regulated in margin
9 - down in expression along longitudinal axis - cell cycle 

## Cluster 10 

```{r}
clusterVis(10)
clusterVis_region(10)
clusterVis_color(10, lcmPaletteColors)
clusterVis_PCA(10) 
clusterVis_region(10)
clusterGO(10)
y <- genesInClust(9, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

```{r}
clusterVis(11)
clusterVis_color(11, lcmPaletteColors)
clusterVis_PCA(11) 
clusterVis_region(10)
clusterGO(11)
y <- genesInClust(11, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```



### Making dataframe for output
```{r}
annotation2 <- read.delim("../../../06diffGeneExp/analysis4_1Sept2017/data/ITAG2.3_all_Arabidopsis_annotated.tsv")

annotation2$ITAG <- gsub("^(.*)[.].*", "\\1", annotation2$ITAG)
annotation2$ITAG <- gsub("^(.*)[.].*", "\\1", annotation2$ITAG)

annotation2 <- annotation2[,c(1:4)]

###Output for DE promoter enrichment.
data.valOutput <- data.val2
names(data.valOutput)[20:21] <- c("cluster", "distanceA")
names(data.valOutput)[1] <- "ITAG"

# subset only genes and cluster number
data.valOutput <- data.valOutput[,c(1, 20)]
## write.csv(data.valOutput, "../data/output/analysis11.large.100917.csv", row.names = FALSE)

## Merge

head(annotation2)
dim(annotation2) #14640
head(data.valOutput)
dim(data.valOutput) #6582

data.valOutput_2 <- merge(data.valOutput, annotation, by = "ITAG",all.x = TRUE )

dim(data.valOutput_2) #6582

# write.csv(data.valOutput_2, "../data/output/analysis11.large.100917_with_annotations.csv", row.names = FALSE)

data.valOutput_2 <- read.csv("../data/output/analysis11.large.100917_with_annotations.csv")
```

## Summary on NA values.  


```{r}
head(data.valOutput_2)

nrow(data.valOutput_2) #6582
sum(is.na(data.valOutput_2$gene_name)) # 4217

4217 / 6582 # 65% of genes in this list are NA.

data.valOutput_2$cluster <- as.character(data.valOutput_2$cluster)

summary_4 <- data.valOutput_2 %>%
  group_by(cluster) %>%
  summarize(sum(is.na(gene_name))) 

summary_5 <- data.valOutput_2 %>%
  group_by(cluster) %>%
  count()

merge(summary_4, summary_5, by = "cluster") 

## Cluster 10
69/ 108 # 63.8 % is NA
57/112 #51.8 %

## Yasu's leaf development genes




data.valOutput_2 %>%
  filter(cluster == "10") %>%
  nrow()

data.valOutput_2 %>%
  filter(cluster == "10" | cluster == "11") %>%
  nrow()
```


### Visualization


```{r}
one <- clusterVis_color(1, lcmPaletteColors) 

numbers <- c(1:36)


for (i in 1:36) {
  plot_name <- paste0("plot_", i) 
  assign(plot_name, clusterVis_color(i, lcmPaletteColors))
}

# make list of plot names
# plots <- paste0("plot_", numbers) 

plot_grid(plot_1, plot_2, plot_3, plot_4, plot_5, plot_6, plot_7, plot_8,
  plot_9, plot_10, plot_11, plot_12, plot_13, plot_14, plot_15, plot_16,
  plot_17, plot_18, plot_19, plot_20, plot_21, plot_22, plot_23, plot_23,
  plot_25, plot_26, plot_27, plot_28, plot_29, plot_30, plot_31, plot_32, 
  plot_33, plot_34, plot_35, plot_36, labels = c(1:36), ncol = 4)

plot_grid()
```


## Cluster 10 and 11 visualization

```{r}

sub_cluster <- subset(data.val2, som$unit.classif == 10 )
sub_data <- sub_cluster[,c(1, 8:13)] # just the sample types

m.data <- melt(sub_data)
m.data$region <- ifelse(grepl("wta", m.data$variable, ignore.case = T), "top", 
                        ifelse(grepl("wtb", m.data$variable, ignore.case = T), "mid", "base"))

m.data$tissue <- ifelse(grepl("other", m.data$variable, ignore.case = T), "rachis", 
                        ifelse(grepl("mbr", m.data$variable, ignore.case = T), "margin", "NA"))

cluster10 <- m.data

sub_cluster <- subset(data.val2, som$unit.classif == 10 )
sub_data <- sub_cluster[,c(1, 8:13)] # just the sample types

m.data <- melt(sub_data)
m.data$region <- ifelse(grepl("wta", m.data$variable, ignore.case = T), "top", 
                        ifelse(grepl("wtb", m.data$variable, ignore.case = T), "mid", "base"))

m.data$tissue <- ifelse(grepl("other", m.data$variable, ignore.case = T), "rachis", 
                        ifelse(grepl("mbr", m.data$variable, ignore.case = T), "margin", "NA"))

cluster11 <- m.data


cluster10$cluster <- 10
cluster11$cluster <- 11

m.data <- rbind(cluster10, cluster11)

p <- ggplot(m.data, aes(y = value, x = tissue, color = tissue))
p + geom_point(alpha = 0.5,position = "jitter", size = 1) + 
  geom_boxplot(alpha = 0.70, outlier.size = 0) +
  scale_colour_manual(values = c( "#bdbec0","#4da2da")) +
  theme_bw(base_size = 25) + 
  facet_grid(region~cluster) +
  labs(x = "tissue", y = "scaled value") 
```


## This seems weird. I think maybe I should make a heatmap to display this?


```{r}
library(superheat)

heat_data <- data.val2[,c(1, 8:13)]

## Make first row, rownames
heat_data2 <- heat_data[,-1]
rownames(heat_data2) <- heat_data[,1]

### Without cluster assignments
superheat(heat_data2, scale = FALSE)

#With SOM cluster assignments
## Still need to order numerically

SOMclusters <- data.val2$unitClassIfA
superheat(heat_data2, scale = FALSE, membership.rows = SOMclusters, order.rows = order(data.val2$unitClassIfA))
```

## Now subsetted

```{r}
subset_heat <- subset(data.val2, unitClassIfA == 10 | unitClassIfA == 11)

heat_data <- subset_heat[,c(1, 8:13)]

## Make first row, rownames
heat_data2 <- heat_data[,-1]
rownames(heat_data2) <- heat_data[,1]

### Without cluster assignments
superheat(heat_data2, scale = FALSE)

#With SOM cluster assignments
## Still need to order numerically

SOMclusters <- subset_heat$unitClassIfA

superheat(heat_data2, scale = FALSE, membership.rows = SOMclusters)
```

## GO analysis of Cluster 10 and 11

No go terms associated with these clusters, even when combined together.
```{r}

##Sets up plot
dev.off()
plot.new()
  
#sub_cluster
sub_cluster <- subset(data.val2, som$unit.classif == 10 | som$unit.classif == 11)

itag.sc <- as.data.frame(sub_cluster$gene) 
  colnames(itag.sc)[1] <- "itag"
  itag.sc$sc <- 1    
 
itag.sc <- unique(itag.sc) #Check. Make sure each itag is represented only once.
  
#Merge all by itag
matrixGO <- merge(itag.sc, geneLength, by = "itag", all = TRUE)
matrixGO[is.na(matrixGO)] <- 0
pat <- matrixGO
  
#Now that we have the data in the right format, we can proceed with GO enrichment.
  
genes = as.integer(pat[,"sc"])
names(genes) = pat$itag
table(genes)
length(genes)
  
pwf = nullp(genes,bias.data = pat$length)
  
GO.wall = goseq(pwf,gene2cat = cate)
head(GO.wall)
  
#This is going to correct for multiple testing.  You can specify the p-value cut-off of GO categories you are interested.
  
enriched.GO = GO.wall$category[p.adjust(GO.wall$over_represented_pvalue, method = "BH") < 0.05]
  
enriched.GO
  
my.GO <- as.character(enriched.GO)
my.GO.table <- Term(my.GO)
my.GO.table
t <- as.matrix(my.GO.table)

print(t)
```