GO Enrichment
-------------

```{r}
library(tidyr)
library(goseq)
library(GO.db)
```

Setting up the DE table for GO analysis
-----------------

### File Input

Input the output from DE analysis.  This is made for a list that includes only the significant genes.  

```{r}
sigOnly <- read.csv("../data/output/02_wt.question1.onlysig_0.05_.txt", header = TRUE)
head(sigOnly)
sigOnly$logFC <- as.numeric(as.character(sigOnly$logFC))
colnames(sigOnly)[1] <- "itag" 
head(sigOnly)
```

### Subset 

First I need to subset the list to up or down regulated, then add a new colum that specififys 1.  This column is need to for merging. 

```{r}
upITAG <- subset(sigOnly, logFC > 0, select = c(itag))
upITAG$up <- 1 
    
downITAG <- subset(sigOnly, logFC < 0, select = c(itag))
downITAG$down <- 1 
  
allITAG <- subset(sigOnly, select = c(itag))
allITAG$all <- 1
```

### Merge I - with normalized ITAG length gene list

read in guide. 
```{r}
geneLength <- read.csv("../data/normalized_genes_length.csv")

## remove trailing numbers in ITAG
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)

#isolate just the gene list
genes <- subset(geneLength, select = c(itag))
```

First merge each table to geneLength
```{r}
head(genes)
dim(genes)
head(upITAG)

upITAGmerge <- merge(genes, upITAG, by = "itag", all = TRUE)
downITAGmerge <- merge(genes, downITAG, by = "itag", all = TRUE)
allITAGmerge <- merge(genes, allITAG, by = "itag", all = TRUE)
```

### Merge II - Merge them all together.

```{r}
matrixGOupdown <- merge(upITAGmerge, downITAGmerge, by = "itag", all = TRUE)
matrixGOupdownall <- merge(matrixGOupdown, allITAG, by = "itag", all = TRUE)
matrixGO <- merge(matrixGOupdownall, geneLength, by = "itag", all = TRUE)
```

###Clean Up
```{r}
matrixGO[is.na(matrixGO)] <- 0
head(matrixGO)

## This is if you want to write out the table of the GO matrix.
# write.table(matrixGO, "mydata.txt", sep="\t", quote= FALSE)
```

GO enrichment
-----------------

The is the input of the GOslim categories.  There are only two columns 1. itag and 2. go

```{r}
pat <- matrixGO
head(pat)

## New GO table
cate <- read.table("../data/ITAG3.2_protein_go.tsv")
colnames(cate) <- c("itag", "go")

summary(cate$itag)

## remove the trailing num in itag id
cate$itag <- gsub("^(.*)[.].*", "\\1", cate$itag)
cate$itag <- gsub("^(.*)[.].*", "\\1", cate$itag)

## Check
head(cate)

## This is to sepererate the GO column into seperate columns for all the genes that are part of several GO categories.  Basically formatting the new GO table.
## There are warning that there are too few categories. This is okay. It is not okay if there are not enough columns. 

cate_2 <- separate(data = cate, col = go, into = c("go1", "go2", "g04", "go5", "go6", "go7", "go8", "go9"), sep = "\\|")

## Check it out
head(cate_2)

cate_3 <- gather(cate_2, itag, go1:go9, factor_key = TRUE)

colnames(cate_3)[3] <- "go"

## You can see here what the gather function did exactly and what needs to be fixed.
test <- subset(cate_3, itag == "Solyc00g005000")

## To Fix
## First remove rows with NA in go
cate_4 <- cate_3[complete.cases(cate_3), ]

## Now every go term and itag pair is represented only once, so we can get rid of itag.1 column
cate_5 <- cate_4[,-2]

cate <- cate_5
```

### Subseting for GO analysis

Specify the column you are interested in `pat$all` refers to all the DE gene regardless if they are up or down regulated. If you want to specify down regulated, specify `pat$down`. I am going to put this into a loop, where each time the loop goes thought it will perform GO enrichment on all three types of lists of significant genes and them write them to a table. 

```{r}
sigType <- c("up", "down", "all")

for (type in sigType) { 
  genes = as.integer(pat[,type])
  names(genes) = pat$itag
  table(genes)
  length(genes)

  pwf = nullp(genes, bias.data = pat$length)

  GO.wall = goseq(pwf, gene2cat = cate)
  head(GO.wall)

#This is going to correct for multiple testing.  You can specify the p-value cut-off of GO categories you are interested.

  enriched.GO = GO.wall$category[p.adjust(GO.wall$over_represented_pvalue, method = "BH") < 0.05]

  enriched.GO

  my.GO <- as.character(enriched.GO)
  my.GO.table <- Term(my.GO)
  my.GO.table
  t <- as.matrix(my.GO.table)

  print(type) #this is for the knitr document
  print(t) #this is for the knitr document

  ## To write out tables
  write.csv(t, paste0("../data/output/DE1_sigonly_GO_", type, ".txt"))
  write.csv(GO.wall, paste0("../data/output/DE1_sigValues_GO_", type, ".txt"))
}

## This is a long way of doing it. But basically I want all the results in a single dataframe.
## So just concatenate together
all <- read.csv("../data/output/DE1_sigonly_GO_all.txt")
all$type <- "all"

up <- read.csv("../data/output/DE1_sigonly_GO_up.txt")
up$type <- "up"

down <- read.csv("../data/output/DE1_sigonly_GO_down.txt")
down$type <- "down"
 
everything <- rbind(down, up, all)
colnames(everything) <- c("GO_ID", "GO_term", "type")
head(everything)

write.csv(everything, "../data/output/wt_modelled_GO.csv")

### Now with Values

all <- read.csv("../data/output/DE1_sigValues_GO_all.txt")
all$type <- "all"

up <- read.csv("../data/output/DE1_sigValues_GO_up.txt")
up$type <- "up"

down <- read.csv("../data/output/DE1_sigValues_GO_down.txt")
down$type <- "down"
 
everything <- rbind(down, up, all)
everything <- everything[,-1]

write.csv(everything, "../data/output/wt_modelled_GO_with_sig_values.csv")
```

