---
output: pdf_document
---
# Analysis 10 - Comparing clusters of Wildtype

Based off of `lcmSOM_analysis10_122916_redo`, with newly mapped genes.

## Purpose 

Comparing the clusters for two purposes

1. Quality of Analysis - To understand how consistent different clusteres are, therefore how conservative the results are.

2. Decrease the amount of interesting genes in clusters by comparing WT SOM with *tf2* superSOM. 

## Analysis Set-up

Required Libraries

```{r}
library(VennDiagram)
library(ggplot2)
library(reshape)
library(kohonen)
library(goseq)
library(GO.db)
library(cowplot)
library(knitr)
source("./clusterFunctions.R")
```

Upload that dataset:

```{r}
genes25 <- read.csv("../data/output/analysis4.top25_19Oct2017.csv")
colnames(genes25)

## Only WT
genes25 <- genes25[,c(2,9:14)]
```

## Initial PCA 
```{r}
scale_data <- as.matrix(t(scale(t(genes25[c(2:7)])))) 
pca <- prcomp(scale_data, scale = TRUE) 
summary(pca) 
pca.scores <- data.frame(pca$x)
data.val <- cbind(genes25, scale_data, pca.scores) 

PCA12 <- ggplot(data.val, aes(PC1, PC2)) +
  geom_point(alpha = .3) +
  theme_bw(base_size = 30)

PCA34 <- ggplot(data.val, aes(PC3, PC4)) +
  geom_point(alpha = .3) +
  theme_bw(base_size = 30)

plot_grid(PCA12, PCA34, labels = c("", ""), ncol = 1)

### With cluster 1
## plot_grid(PCA12, p1, labels = c("", ""), ncol = 1)
```

## SOM 
```{r}
set.seed(6)
som <- som(scale_data, somgrid(3,2)) # This is where you change the size of the map
summary(som)

lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
lcmPalette <- function(n, alpha =1){
  lcmPaletteColors[1:n]
}
  
plot(som, type = "codes", palette.name = lcmPalette)

## Bind information
data.val2 <- cbind(data.val,som$unit.classif,som$distances) 
names(data.val2)
##rename 
names(data.val2)[20:21] <- c("unitClassIfA", "distanceA")
```

Read in geneLength and cate to run clusterGO function.

```{r}
geneLength <- read.csv("../../../07GO_enrichment/requisiteData/normalized_genes_length.csv")

# Fix itag ids
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)

cate <- read.csv("../../../07GO_enrichment/requisiteData/ITAG3.2_protein_go_clean.csv", header = TRUE)

# Palette
lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
```

##Cluster 1
```{r}
clusterVis(1)
clusterVis_region(1)
clusterVis_color(1, lcmPaletteColors)
p1 <- clusterVis_PCA(1) 
clusterGO(1)
y <- genesInClust(1, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 2
```{r}
clusterVis(2)
clusterVis_region(2)
clusterVis_color(2, lcmPaletteColors)
clusterVis_PCA(2)
clusterGO(2)
y <- genesInClust(2, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```


## Cluster 3

```{r}
clusterVis(3)
clusterVis_region(3)
clusterVis_color(3, lcmPaletteColors)
clusterVis_PCA(3)
clusterGO(3)
y <- genesInClust(3, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 4

```{r}
clusterVis(4)
clusterVis_region(4)
clusterVis_color(4, lcmPaletteColors)
clusterVis_PCA(4)
clusterGO(4)
y <- genesInClust(4, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

##Cluster 5

```{r}
clusterVis(5)
clusterVis_region(5)
clusterVis_color(5, lcmPaletteColors)
clusterVis_PCA(5)
clusterGO(5)
y <- genesInClust(5, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```


##Cluster 6

```{r}
clusterVis(6)
clusterVis_region(6)
clusterVis_color(6, lcmPaletteColors)
clusterVis_PCA(6)
clusterGO(6)
y <- genesInClust(6, data.val2, annotation)
kable(y, format = "latex", booktabs = TRUE)
```

## Superheat

## This seems weird. I think maybe I should make a heatmap to display this?


```{r}
library(superheat)

heat_data <- data.val2[,c(1, 8:13)]

## Make first row, rownames
heat_data2 <- heat_data[,-1]
rownames(heat_data2) <- heat_data[,1]

head(heat_data)
### Without cluster assignments
superheat(heat_data2, scale = FALSE)

#With SOM cluster assignments
## Still need to order numerically

SOMclusters <- data.val2$unitClassIfA
superheat(heat_data2, scale = FALSE, membership.rows = SOMclusters, order.rows = order(data.val2$unitClassIfA))
```

## Now subsetted

```{r}
subset_heat <- subset(data.val2, unitClassIfA == 10 | unitClassIfA == 11)

heat_data <- subset_heat[,c(1, 8:13)]

## Make first row, rownames
heat_data2 <- heat_data[,-1]
rownames(heat_data2) <- heat_data[,1]

### Without cluster assignments
superheat(heat_data2, scale = FALSE)

#With SOM cluster assignments
## Still need to order numerically

SOMclusters <- subset_heat$unitClassIfA

superheat(heat_data2, scale = FALSE, membership.rows = SOMclusters)






### Making dataframe for output
```{r}
###Output for DE promoter enrichment.
data.valOutput <- data.val2
names(data.valOutput)[20:21] <- c("cluster", "distanceA")
names(data.valOutput)[1] <- "ITAG"

# subset only genes and cluster number
data.valOutput <- data.valOutput[,c(1, 20)]
write.csv(data.valOutput, "../data/output/analysis10.small.100917.csv", row.names = FALSE)
```

## Visualizing results of GO enrichment

First I need to get a dataframe that describes the 
```{r}

## GO enrichment of each cluster (based off clusterGO() function)

results <- data.frame()

for(clustNum in 1:6){
  #sub_cluster
  sub_cluster <- subset(data.val2, som$unit.classif == clustNum)
  
  itag.sc <- as.data.frame(sub_cluster$gene) 
  colnames(itag.sc)[1] <- "itag"
  itag.sc$sc <- 1    

  itag.sc <- unique(itag.sc) #Check. Should cut the list in half. # dim(itag.sc) before and after
  
  #Merge all by itag
  matrixGO <- merge(itag.sc, geneLength, by = "itag", all = TRUE)
  matrixGO[is.na(matrixGO)] <- 0
  pat <- matrixGO
  
  #Now that we have the data in the right format, we can proceed with GO enrichment.
  
    genes = as.integer(pat[,"sc"])
    names(genes) = pat$itag
    table(genes)
    length(genes)
  
    pwf = nullp(genes,bias.data=pat$length)
  
    GO.wall = goseq(pwf,gene2cat = cate)
    
    GO.wall$cluster <- clustNum
    results <- rbind(results, GO.wall)
}

# Check results
dim(results)
head(results)
results$cluster <- as.factor(results$cluster)
levels(results$cluster)
summary(results$cluster)

#Adjust P-value
results$BHadjusted <- p.adjust(results$over_represented_pvalue, method = "BH") 
head(results)

```

### Visualization

```{r}
#subset only significant

# subset only significantly enriched
results_sig <- subset(results, BHadjusted < .05)
dim(results_sig) # 66 GO terms
levels(as.factor(results_sig$term)) #47 GO terms

# Write out for supplemental dataset

## write.csv(results_sig, "../data/output/lcmSOM_analysis10_100917_redo_sigGOterms.csv")
```

```{r}
colnames(results_sig)

#remove NA terms
results_sig <- results_sig[!is.na(results_sig$term),]

#clean name so they are shorter
results_sig$term <- gsub("transcription factor activity, sequence-specific DNA binding", "transcription factor activity", results_sig$term)

results_sig$term <- gsub("positive regulation of transcription from RNA polymerase II promoter", "transcription from RNA pol II", results_sig$term)

results_sig$term <- gsub("glucan endo-1,3-beta-glucanase activity, C-3 substituted reducing group" , "glucan endo-1,3-beta-glucanase activity", results_sig$term)

results_sig$term <- gsub("hydrolase activity, hydrolyzing O-glycosyl compounds", "hydrolase activity", results_sig$term)

results_sig$term <- gsub("induced systemic resistance, jasmonic acid mediated signaling pathway", "induced systemic resistance by jasmonic acid", results_sig$term)

results_sig$term <- gsub("oxidoreductase activity, acting on paired donors, with incorporation or reduction of molecular oxygen", "oxidoreductase activity", results_sig$term)

results_sig$term <- gsub("RNA polymerase II regulatory region sequence-specific DNA binding", "RNA polymerase II seq-specific DNA binding" , results_sig$term)

results_sig$term <- gsub("transferase activity, transferring hexosyl groups" , "transferase activity", results_sig$term)

levels(as.factor(results_sig$term))

# Remove redundant GO terms

remove_terms <- c("transport", "chitin catabolic process", "drug transmembrane transporter activity", "induced systemic resistance by jasmonic acid","endopeptidase inhibitor activity" )


results_sig <- subset(results_sig, !(results_sig$term %in% remove_terms))

levels(as.factor(results_sig$term))

## For Ordering terms in Visualization
results_sig <- with(results_sig, results_sig[order(cluster),])
results_sig$term <- factor(results_sig$term, levels = unique(as.character(results_sig$term)) )

ggplot(results_sig, aes(cluster, term, size = BHadjusted)) +
  geom_point(alpha = .9) +
  scale_size(range = c(1, 5)) +
  theme_bw() +
  theme(# axis.text.x = element_blank(),
        # axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 19)) +
  labs(size="p-value")

```


```{r}

one <- clusterVis_color(1, lcmPaletteColors) 
two <- clusterVis_color(2, lcmPaletteColors)
three <- clusterVis_color(3, lcmPaletteColors)
four <- clusterVis_color(4, lcmPaletteColors)
five <- clusterVis_color(5, lcmPaletteColors)
six <- clusterVis_color(6, lcmPaletteColors)

plot_grid(one, two, three, four, five, six, labels = c("1", "2", "3", "4", "5", "6"), ncol = 2)
```