---
editor_options:
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---
# Super SOM - Large

## Purpose

The purpose of this analysis is to make a superSOM. This is a 6 x 6 hexogonal plot.

```{r}
library(ggplot2)
library(reshape)
library(plyr)
library(kohonen)
source("../r/clusterFunctions.R")
```

## PCA
Upload that dataset:

```{r}
genes25 <- read.csv("../data/output/analysis4.top25_19Oct2017.csv")

genes25 <- genes25[,c(2:14)]
m.genes25 <- melt(genes25)
# head(m.genes25)
  
names(m.genes25) <- c("gene", "sample", "mean")
  
#set genotype
  
m.genes25$genotype <- ifelse(grepl("wt", m.genes25$sample, ignore.case = T), "wt", 
           ifelse(grepl("tf2", m.genes25$sample, ignore.case = T), "tf2", "unknown"))
  
#set tissue
  
m.genes25$tissue <- ifelse(grepl("other", m.genes25$sample, ignore.case = T), "other", 
           ifelse(grepl("mbr", m.genes25$sample, ignore.case = T), "mbr", "unknown"))
  
  #Set Region
m.genes25$region <- ifelse(grepl("a", m.genes25$sample, ignore.case = T), "A", 
           ifelse(grepl("c", m.genes25$sample, ignore.case = T), "C", "B"))
  
#Set type
  
m.genes25$type <- paste(m.genes25$region, m.genes25$tissue,  sep = "")
  
m.genes25.sub <- m.genes25[,c(1,7,4,3)]
# head(m.genes25.sub)
  
#Change from long to wide data format
m.genes25.long <- cast(m.genes25.sub, genotype + gene ~ type, value.var = mean, fun.aggregate = "mean")  
m.genes25.long <- as.data.frame(m.genes25.long) 
```
  
## Scaling the Data seperately 

```{r}
# head(m.genes25.long)
wt <- subset(m.genes25.long, genotype == "wt")
tf2 <- subset(m.genes25.long, genotype == "tf2")
  
#transformation.
scale_data.wt <- as.matrix(t(scale(t(wt[c(3:8)]))))
scale_data.tf2 <- as.matrix(t(scale(t(tf2[c(3:8)]))))

scale_data_sep <- rbind(scale_data.tf2,scale_data.wt)
```
  
Continuing on with PCA
```{r}
pca_sep <- prcomp(scale_data_sep) 
  
summary(pca_sep) 
  
pca.scores_sep <- data.frame(pca_sep$x)
  
data.val_sep <- cbind(m.genes25.long, scale_data_sep, pca.scores_sep) 
```
  
Visualizing the PCA
  
```{r}
p <- ggplot(data.val_sep, aes(PC1, PC2, color = genotype))
p + geom_point(alpha = 0.5)
```
  
## Scaling the data together
  
```{r}
#transformation
scale_data_tog <- as.matrix(t(scale(t(m.genes25.long[c(3:8)]))))
```
  
  
Continuing on with PCA
```{r}
pca_tog <- prcomp(scale_data_tog) 
  
summary(pca_tog) 
  
pca.scores_tog <- data.frame(pca_tog$x)
  
data.val_tog <- cbind(m.genes25.long, scale_data_tog, pca.scores_tog) 
  
```
  
  
## Visualizing the PCA

What is the difference if you scale the genes seperately?
  
```{r}
p <- ggplot(data.val_tog, aes(PC1, PC2, color = genotype)) 
p + geom_point(alpha = 0.5)

p <- ggplot(data.val_sep, aes(PC1, PC2, color = genotype)) 
p + geom_point(alpha = 0.5)
```
  
This doesn't seem to affect the analysis very much. 

## SuperSOM
  
```{r}
## Using the the version where the values were scaled seperatley.
# head(data.val_sep)
data.val <- data.val_sep
  
set.seed(6)
names(data.val)
# head(data.val)

## Isolate only the scaled values as matrices
tf2 <- as.matrix(subset(data.val, genotype == "tf2", select = 9:14))
wt <- as.matrix(subset(data.val, genotype == "wt", select = 9:14))
 
# Make sure they are in proper order
all.data <- list(tf2, wt)
# head(all.data)
```

## SOM

```{r}
## Making the SOM map
ssom <- supersom(all.data, somgrid(6, 6, "hexagonal")) 
  
summary(ssom)

#par(mfrow = c(3, 2))
plot(ssom, type = "changes")
plot(ssom, type = "codes")
plot(ssom, type = "counts")
plot(ssom, type = "quality")
  
data.val <- cbind(data.val,ssom$unit.classif,ssom$distances)
  
# head(data.val)
  
## write.table(data.val, file = "../data/output/ssom.data.analysis5c_05Nov2017_large.txt")
```

### Visualization 

```{r}
## Read in Data from previous section
plot.data <- read.table("../data/output/ssom.data.analysis5c_05Nov2017_large.txt", header = TRUE)
names(plot.data)
dim(plot.data)

## Princliple components colored by clusters
p <- ggplot(plot.data, aes(PC1, PC2, colour = factor(ssom.unit.classif))) 
p + geom_point(alpha = .5) + 
  theme_bw() +
  facet_grid(.~genotype)
```


## Each of the clusters

```{r}
data.val2 <- read.table("../data/output/ssom.data.analysis5c_05Nov2017_large.txt", header = TRUE)

clusterVis_region_ssom(1)
clusterVis_line_ssom(1)

clusterVis_region_ssom(2)
clusterVis_line_ssom(2)

clusterVis_region_ssom(3)
clusterVis_line_ssom(3)

clusterVis_region_ssom(4)
clusterVis_line_ssom(4)

clusterVis_region_ssom(5)
clusterVis_line_ssom(5)

clusterVis_region_ssom(6)
clusterVis_line_ssom(6)

clusterVis_region_ssom(7)
clusterVis_line_ssom(7)

clusterVis_region_ssom(8)
clusterVis_line_ssom(8)

clusterVis_region_ssom(9)
clusterVis_line_ssom(9)

clusterVis_region_ssom(10)
clusterVis_line_ssom(10)

clusterVis_region_ssom(11)
clusterVis_line_ssom(11)

clusterVis_region_ssom(12)
clusterVis_line_ssom(12)

clusterVis_region_ssom(13)
clusterVis_line_ssom(13)

clusterVis_region_ssom(14)
clusterVis_line_ssom(14)

clusterVis_region_ssom(15)
clusterVis_line_ssom(15)

clusterVis_region_ssom(16)
clusterVis_line_ssom(16)

clusterVis_region_ssom(17)
clusterVis_line_ssom(17)

clusterVis_region_ssom(18)
clusterVis_line_ssom(18)

clusterVis_region_ssom(19)
clusterVis_line_ssom(19)

clusterVis_region_ssom(20)
clusterVis_line_ssom(20)

clusterVis_region_ssom(21)
clusterVis_line_ssom(21)

clusterVis_region_ssom(22)
clusterVis_line_ssom(22)

clusterVis_region_ssom(23)
clusterVis_line_ssom(23)

clusterVis_region_ssom(24)
clusterVis_line_ssom(24)

clusterVis_region_ssom(25)
clusterVis_line_ssom(25)

clusterVis_region_ssom(26)
clusterVis_line_ssom(26)

clusterVis_region_ssom(27)
clusterVis_line_ssom(27)

clusterVis_region_ssom(28)
clusterVis_line_ssom(28)

clusterVis_region_ssom(29)
clusterVis_line_ssom(29)

clusterVis_region_ssom(30)
clusterVis_line_ssom(30)

clusterVis_region_ssom(31)
clusterVis_line_ssom(31)

clusterVis_region_ssom(32)
clusterVis_line_ssom(32)

clusterVis_region_ssom(33)
clusterVis_line_ssom(33)

clusterVis_region_ssom(34)
clusterVis_line_ssom(34)

clusterVis_region_ssom(35)
clusterVis_line_ssom(35)

clusterVis_region_ssom(36)
clusterVis_line_ssom(36)
```

Make data table with gene names
```{r}
#Prereq annotation files for function
annotation1<- read.delim("../../../06diffGeneExp/analysis1_2014/DE/DE_analysis/beforeAnalysis/requisiteData/ITAG2.3_all_Arabidopsis_ITAG_annotations.tsv", header=FALSE)  #Changed to the SGN human readable annotation
colnames(annotation1) <- c("ITAG", "SGN_annotation")
annotation2<- read.delim ("../../../06diffGeneExp/analysis1_2014//DE/DE_analysis/beforeAnalysis/requisiteData/ITAG2.3_all_Arabidopsis_annotated.tsv")
annotation <- merge(annotation1,annotation2, by = "ITAG")

#Only Gene Name and ITAG
names(annotation)
annotation <- annotation[,c(1,4,5)]

#fix to one regex
annotation$ITAG <- gsub("^(.*)[.].*", "\\1", annotation$ITAG)
annotation$ITAG <- gsub("^(.*)[.].*", "\\1", annotation$ITAG)

data.val3 <- data.val2
colnames(data.val3)[2] <- "ITAG"
everything <- merge(data.val3, annotation, by = "ITAG", all.x = TRUE)

write.csv(everything, file = "../data/output/ssom.data.analysis5c_36_29March2018_geneList.csv")
```