---
output: html_document
editor_options: 
  chunk_output_type: console
---
 ---
output: pdf_document
---
# Question 1 - Wildtype

1. **Are there genes that are differentially expressed between MBR and Rachis in all three groups across the longitudinal axis?** Genotype specific. Basically asking what are the genes that define the Marginal Blastozone and what are the genes that define the Rachis.  This analysis is wild type specific. 

## Key to Samples

genotype: either wildtype of *tf2*

region: A. tip B. early emmerging leaflet C. base

type: MBR = Marginal Blastozone Region, other = the rachis or midvein region

## Bring in results from normalization.

```{r, echo=FALSE}
library(reshape2)
# library(mixOmics)

# --------------------- Read in Data

# To re-run normalization
source("01_normalization.R") # Takes about a thirty seconds to run. 
## There are warnings

# To read in output from normalization.R
y <- read.csv("../data/output/normalizedReadCount04March2019.csv", row.names = 1)

designTable <- read.csv("../data/output/designTable_from_01_normalization.R_03March2020.csv", row.names = 1)
```

## Analysis - Wildtype.  

In order to answer the question: Are there genes that are differentially expressed between MBR and Rachis tissue in all three groups across the longitudinal axis? I am going to use an additive linear model, with the longitudinal axis as the blocking factor. This should be genotype specific.

```{r}
## First I need to subset based on genotype
colnames(y)
wtY <- y[,c(1,23:44)]

## This comes from normalization.R
designTable

designTableWT <- designTable[22:43,]

designTableWT
wtRegion <- designTableWT$region
wtTissue <- designTableWT$tissue

design <- model.matrix(~wtRegion + wtTissue)

rownames(design) <- colnames(wtY)

design

head(wtY)
wtY <- wtY %>% mutate_if(is.numeric, list(~na_if(., Inf))) %>% 
  mutate_if(is.numeric, list(~na_if(., -Inf)))

wtY_DGE <- DGEList(counts = wtY[,-1], group=designTableWT$sample)
wtY_DGE <- estimateDisp(wtY_DGE, design, robust = TRUE) #Estimate Dispersion
wtY_DGE$common.dispersion # 0.1488789
```

### WT- Differential Gene Expression

```{r}
fit <- glmFit(wtY_DGE, design)
lrt <- glmLRT(fit)
```

Here we see the top tags for MBR vs rachis tissue differential. It is adjusted for baseline differences between top, middle, and base. 

```{r}
o <- order(lrt$table$PValue)

cpm(wtY)[o[1:10],] #check to see differences
```

```{r}
summary(de <- decideTestsDGE(lrt))
detags <- rownames(y)[as.logical(de)]
plotSmear(lrt, de.tags = detags)

## abline(h = c(-1, 1), col = "blue")
```

All gene output
```{r}
results <- topTags(lrt, n = Inf)

dim(results$table)
summary(de <- decideTestsDGE(lrt))

sum(results$table$FDR < .05) # How many are DE genes total? #816 or #2015

#Subset only significant
results.sig <- subset(results$table, results$table$FDR < 0.05) 

sig.genes <- results.sig$genes #only gene names
length(sig.genes)
```

What are these genes?
For this we need to add some annotation.

Essentially we are merging two annotations files to 1.) only sig genes 2.) all genes

```{r}
annotation1 <- read.delim("../data/ITAG2.3_all_Arabidopsis_ITAG_annotations.tsv", header = FALSE)  #Changed to the SGN human readable annotation
colnames(annotation1) <- c("ITAG", "SGN_annotation")
annotation2 <- read.delim("../data/ITAG2.3_all_Arabidopsis_annotated.tsv")
annotation <- merge(annotation1,annotation2, by = "ITAG")

## head(annotation)
colnames(annotation)
colnames(results.sig)
## head(results.sig)

annotation$ITAG <- gsub("^(.*)[.].*", "\\1", annotation$ITAG)
annotation$ITAG <- gsub("^(.*)[.].*", "\\1", annotation$ITAG)


#Change headers for merging
colnames(results.sig)[1] <- "ITAG"
colnames(annotation)[1] <- "ITAG"
colnames(results$table)[1] <- "ITAG"

results.sig.annotated <- merge(results.sig, annotation, by = "ITAG", all.x = TRUE) #This is merging only sig genes

dim(results.sig.annotated)
head(results.sig.annotated)

#Making all table
head(results$table)
results.all.annotated <- merge(results$table, annotation, by = "ITAG")

#Write out table to file
## write.table(results.all.annotated, file = "../data/output/wt.question1_all_08January2018.txt", sep = "\t",row.names = F) 

## write.csv(results.sig.annotated, file = "../data/output/wt.question1.onlysig_0.05_.txt", row.names = F) 
```

