---
title: "Flow Cytometry Analysis"
output: pdf_document
---

What do I need to make sure I illustrate:

1. In tomato leaves there is a slight change from 2C to 4C, but not much beyond that.

2. There is no difference between tip and base in plants.

Read in data and establish dependencies.
```{r}
library(gdata)
library(plyr)
library(ggplot2)
library(reshape2)

flowC = read.xls("../data/flowCytometry.xlsx", sheet = 1, header = TRUE)
names(flowC)
str(flowC)
dim(flowC)
```

Reshape data.

```{r}
#subset only what I need
names(flowC)
levels(flowC$section)

#fix typo
flowC$section <- gsub("tip ", "tip", flowC$section)

# Clean up
flowCSub1 <- flowC[,c(1:6, 10:12, 15)]
names(flowCSub1)
str(flowCSub1)
flowCSub1[is.na(flowCSub1)] <- 0
flowCMelt <- melt(flowCSub1, id.vars = c("leafSample", "genotype", "plant", "leaf", "section", "area", "age"))

colnames(flowCMelt)[8] <- "peaks"
colnames(flowCMelt)[9] <- "flowValue"
```

## Genotype

We are not using the entire-2 data.  Only wildtype
There are two data sets 1. flowCMelt and 2.flowCSub1, make for wt only.

```{r}
flowCMelt <- subset(flowCMelt, genotype == "wt")
flowCSub1 <- subset(flowCSub1, genotype == "wt")
```


## Age
#Since age of (30 and 31) and (90 and 91) are essentially the same, I am going to collapse them. 

```{r}
flowCMelt$age <- as.factor(flowCMelt$age) #treat age as factor
flowCMelt$age <- sub("31", "30", flowCMelt$age)
flowCMelt$age <- sub("91", "60", flowCMelt$age)

flowCMelt.summary <- ddply(flowCMelt, c("genotype", "peaks","age"), summarise,
            N  = length(flowValue), 
            flowCyt = mean(flowValue),
            sd = sd(flowValue),
            se = sd / sqrt(N) )
                    
flowCMelt.summary

#Only WT

wtOnly <- subset(flowCMelt.summary, genotype == "wt")
```

## Visualize 1

First visualize across age. 

```{r}
## Change Data for Plot 
wtOnly$peaks <- gsub("percent", "", wtOnly$peaks)
wtOnly$age <- paste(wtOnly$age,"days", sep=" ")

## change order of age, so facetting is correct

wtOnly$age <- factor(wtOnly$age,
         levels = c("8 days", "18 days", "30 days", "60 days", "90 days"))

## Change decimal places
wtOnly[,c(5:7)] <- wtOnly[,c(5:7)] * 100

# Plot with error bars
ggplot(wtOnly, aes(genotype, flowCyt, fill = peaks)) +
   geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin=flowCyt - se, ymax=flowCyt + se), 
                width=.2, 
                colour="black", 
                position = position_dodge(.9)) +
  theme_bw() +
  facet_grid(.~age) +
  scale_fill_brewer() +
  xlab("DNA content peaks") +
  ylab("% of nuclei") +
  theme(strip.background = element_rect(color = "000000", fill="#FFFFFF"),
        text = element_text(size = 18),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

## Visualize 2

Look at differences between tip and base.

```{r}
#I only want to use one sample per leaf
#remove cotolydons
flowCMelt2 <- subset(flowCMelt, section == "tip" | section == "base")

flowCMelt.summary2 <- ddply(flowCMelt2, c("genotype", "age", "section", "peaks"), summarise,
                    N  = length(flowValue),
                    flowCyt = mean(flowValue),
                    sd = sd(flowValue),
                    se = sd / sqrt(N) )

## Only wildtype
flowCMelt.summary2.wt <- subset(flowCMelt.summary2, genotype == "wt")

# Remove 30 days

flowCMelt.summary2.wt <- subset(flowCMelt.summary2.wt, age != 30)

# change wording
flowCMelt.summary2.wt$peaks <- gsub("percent", "", flowCMelt.summary2.wt$peaks)
flowCMelt.summary2.wt$age <- paste(flowCMelt.summary2.wt$age,"days", sep=" ")

## Change decimal places

colnames(flowCMelt.summary2.wt)
flowCMelt.summary2.wt[,c(6:8)] <- flowCMelt.summary2.wt[,c(6:8)] * 100

## Plot

ggplot(flowCMelt.summary2.wt, aes(peaks, flowCyt, fill = section)) +
   geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin=flowCyt - se, ymax=flowCyt + se), 
                width=.2, 
                colour="black", 
                position = position_dodge(.9)) +
  theme_bw() +
  facet_grid(.~age) +
  xlab("DNA content peaks") +
  ylab("percent of nuclei") +
  scale_fill_manual(values=c("black","grey")) +
  theme(strip.background = element_rect(color = "000000", fill="#FFFFFF"),
        strip.text.x = element_text(),
        text = element_text(size = 20))
```




