GO Enrichment
-------------


```{r}
## Libraries
library(tidyr)
library(goseq)
library(GO.db)
library(yaml)
library(rmarkdown)

## Read in YAML guide
### Set Working Directory
rstudioapi::getActiveDocumentContext
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

## Read in sample names from yaml
yamls <- yaml.load_file("de.yml")
sample1 <- yamls$sample1
sample2 <- yamls$sample2

sample1
sample2
```

```{r, eval = FALSE}
## Render
render("GO_basedOnSkeletonGO.Rmd", "pdf_document", output_file = paste(sample1,"_",sample2,"_","GO.pdf",sep = ""))
```

Setting up the DE table for GO analysis
-----------------

### File Input

Input the output from DE analysis.  This is made for a list that includes only the significant genes.  

```{r}
sigOnly <- read.table(paste("../../../../requisiteData/data_06Sept2017/", sample1,"_",sample2,"_DE_sig.txt", sep = ""), header = TRUE, fill = TRUE)
sigOnly$logFC <- as.numeric(as.character(sigOnly$logFC))
colnames(sigOnly)[1] <- "itag" 
```

### Subset 

First I need to subset the list to up or down regulated, then add a new colum that specififys 1.  This column is need to for merging. 

```{r}
upITAG <- subset(sigOnly, logFC > 0, select = c(itag))
upITAG$up <- 1 
    
downITAG <- subset(sigOnly, logFC < 0, select = c(itag))
downITAG$down <- 1 
  
allITAG <- subset(sigOnly, select = c(itag))
allITAG$all <- 1
```

### Merge I - with normalized ITAG length gene list

read in guide. 
```{r}
geneLength <- read.csv("../../../../requisiteData/normalized_genes_length.csv")

## remove trailing numbers in ITAG
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)
geneLength$itag <- gsub("^(.*)[.].*", "\\1", geneLength$itag)

#isolate just the gene list
genes <- subset(geneLength, select = c(itag))
```

First merge each table to geneLength
```{r}

upITAGmerge <- merge(genes, upITAG, by = "itag", all = TRUE)
downITAGmerge <- merge(genes, downITAG, by = "itag", all = TRUE)
allITAGmerge <- merge(genes, allITAG, by = "itag", all = TRUE)
```

### Merge II - Merge them all together.

```{r}
matrixGOupdown <- merge(upITAGmerge, downITAGmerge, by = "itag", all = TRUE)
matrixGOupdownall <- merge(matrixGOupdown, allITAG, by = "itag", all = TRUE)
matrixGO <- merge(matrixGOupdownall, geneLength, by = "itag", all = TRUE)
```

### Clean Up
```{r}
matrixGO[is.na(matrixGO)] <- 0
head(matrixGO)

## This is if you want to write out the table of the GO matrix.
# write.table(matrixGO, "mydata.txt", sep="\t", quote= FALSE)
```

GO enrichment
-----------------

The is the input of the GOslim categories.  There are only two columns 1. itag and 2. go

```{r}
pat <- matrixGO
head(pat)

## New GO table
cate <- read.table("../../requisiteData/ITAG3.2_protein_go.tsv")
colnames(cate) <- c("itag", "go")

summary(cate$itag)

## remove the trailing num in itag id
cate$itag <- gsub("^(.*)[.].*", "\\1", cate$itag)
cate$itag <- gsub("^(.*)[.].*", "\\1", cate$itag)

cate <- separate(data = cate, col = go, into = c("go1", "go2", "g04", "go5", "go6", "go7", "go8", "go9"), sep = "\\|")

cate <- gather(cate, itag, go1:go9, factor_key = TRUE)
colnames(cate)[3] <- "go"

## First remove rows with NA in go
cate <- cate[complete.cases(cate), ]

## Now every go term and itag pair is represented only once, so we can get rid of itag.1 column
cate <- cate[,-2]
head(cate)

## Write out cleaned
## write.csv(cate, "../../requisiteData/ITAG3.2_protein_go_clean.csv", row.names = FALSE)
```

### Subseting for GO analysis

Specify the column you are interested in `pat$all` refers to all the DE gene regardless if they are up or down regulated. If you want to specify down regulated, specify `pat$down`. I am going to put this into a loop, where each time the loop goes thought it will perform GO enrichment on all three types of lists of significant genes and them write them to a table. 

```{r}
sigType <- c("up", "down", "all")

for (type in sigType) { 
  genes = as.integer(pat[,type])
  names(genes) = pat$itag
  table(genes)
  length(genes)

  pwf = nullp(genes, bias.data = pat$length)

  GO.wall = goseq(pwf, gene2cat = cate)
  head(GO.wall)

#This is going to correct for multiple testing.  You can specify the p-value cut-off of GO categories you are interested.

  enriched.GO = GO.wall$category[p.adjust(GO.wall$over_represented_pvalue, method = "BH") < 0.05]

  enriched.GO

  my.GO <- as.character(enriched.GO)
  my.GO.table <- Term(my.GO)
  my.GO.table
  t <- as.matrix(my.GO.table)

  print(type) #this is for the knitr document
  print(t) #this is for the knitr document

  write.table(t, file = paste(sample1,"_",sample2,"DE1_sigonly_",type,"_GO.txt", sep = ""))
write.table(GO.wall, file = paste(sample1,"_",sample2,"DE1_sigValues_",type,"_GO.txt", sep = ""))
  }


```

