---
output: pdf_document
---
# Analysis 11 - Comparing clusters of tf2 large

Based off of `lcmSOM_analysis10_122916_redo`, with newly mapped genes.

## Purpose 

Comparing the clusters for two purposes

1. Quality of Analysis - To understand how consistent different clusteres are, therefore how conservative the results are.

2. Decrease the amount of interesting genes in clusters by comparing WT SOM with *tf2* superSOM. 

## Analysis Set-up

Required Libraries

```{r}
library(ggplot2)
library(reshape)
library(kohonen)
library(goseq)
library(GO.db)
library(knitr)
library(somtools)
library(cowplot)

#source("./clusterFunctions.R")
```

Upload that dataset:

```{r}
genes25 <- read.csv("../data/output/analysis4.top25_19Oct2017.csv")
colnames(genes25)

## Only tf2
genes25 <- genes25[,c(2:8)]
```

```{r}
scale_data <- as.matrix(t(scale(t(genes25[c(2:7)])))) 
pca <- prcomp(scale_data, scale = TRUE) 
summary(pca) 
pca.scores <- data.frame(pca$x)
data.val <- cbind(genes25, scale_data, pca.scores) 

PCA12 <- ggplot(data.val, aes(PC1, PC2)) +
  geom_point(alpha = .3)

PCA34 <- ggplot(data.val, aes(PC3, PC4)) +
  geom_point(alpha = .3)

plot_grid(PCA12, PCA34, labels = c("", ""), ncol = 2)
```

## SOM 

```{r}
# subset only the scaled gene expression values
# colnames(scale_data) <- NULL

set.seed(6)
som <- som(scale_data, somgrid(6,6)) # This is where you change the size of the map
summary(som)

lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
lcmPalette <- function(n, alpha =1){
  lcmPaletteColors[1:n]
}
  
plot(som, type = "codes", palette.name = lcmPalette)

## Bind information
data.val2 <- cbind(data.val,som$unit.classif,som$distances) 
names(data.val2)
##rename 
names(data.val2)[20:21] <- c("unitClassIfA", "distanceA")
```

## Visualization

```{r}
# Palette
lcmPaletteColors <- c( "#b3a2ce", "#4753a4","#bf9e71", "#956025","#b2d9a6", "#0f7c3e")
```

## Cluster 1
```{r}
clusterVis(1)
## clusterVis_region(1)
clusterVis_color(1)
clusterVis_PCA(1) 
## y <- genesInClust(1, data.val2, annotation)
## kable(y, format = "latex", booktabs = TRUE)
```

## Cluster 2
```{r}
clusterVis(2)
clusterVis_color(2)
clusterVis_PCA(2) 
```

## Cluster 3
```{r}
clusterVis(3)
clusterVis_color(3)
clusterVis_PCA(3) 
```

## Cluster 4
```{r}
clusterVis(4)
clusterVis_color(4)
clusterVis_PCA(4) 
```

## Cluster 5
```{r}
clusterVis(5)
clusterVis_color(5)
clusterVis_PCA(5) 
```

## Cluster 6
```{r}
clusterVis(6)
clusterVis_color(6)
clusterVis_PCA(6) 
```

## Cluster 7
```{r}
clusterVis(7)
clusterVis_color(7)
clusterVis_PCA(7) 
```

## Cluster 8
```{r}
clusterVis(8)
clusterVis_color(8)
clusterVis_PCA(8) 
```

## Cluster 9
```{r}
clusterVis(9)
clusterVis_color(9)
clusterVis_PCA(9) 
```

## Cluster 10
```{r}
clusterVis(10)
clusterVis_color(10)
clusterVis_PCA(10) 
```

## Cluster 11
```{r}
clusterVis(11)
clusterVis_color(11)
clusterVis_PCA(11) 
```

## Cluster 12
```{r}
clusterVis(12)
clusterVis_color(12)
clusterVis_PCA(12) 
```

## Cluster 13
```{r}
clusterVis(13)
clusterVis_color(13)
clusterVis_PCA(13) 
```

## Cluster 14
```{r}
clusterVis(14)
clusterVis_color(14)
clusterVis_PCA(14) 
```

## Cluster 15
```{r}
clusterVis(15)
clusterVis_color(15)
clusterVis_PCA(15) 
```

## Cluster 16
```{r}
clusterVis(16)
clusterVis_color(16)
clusterVis_PCA(16)
```

## Cluster 17
```{r}
clusterVis(17)
clusterVis_color(17)
clusterVis_PCA(17) 
```

## Cluster 18
```{r}
clusterVis(18)
clusterVis_color(18)
clusterVis_PCA(18) 
```

## Cluster 19
```{r}
clusterVis(19)
clusterVis_color(19)
clusterVis_PCA(19) 
```

## Cluster 20
```{r}
clusterVis(20)
clusterVis_color(20)
clusterVis_PCA(20) 
```

## Cluster 21
```{r}
clusterVis(21)
clusterVis_color(21)
clusterVis_PCA(21) 
```

## Cluster 22
```{r}
clusterVis(22)
clusterVis_color(22)
clusterVis_PCA(22) 
```

## Cluster 23
```{r}
clusterVis(23)
clusterVis_color(23)
clusterVis_PCA(23) 
```

## Cluster 24
```{r}
clusterVis(24)
clusterVis_color(24)
clusterVis_PCA(24) 
```

## Cluster 25
```{r}
clusterVis(25)
clusterVis_color(25)
clusterVis_PCA(25) 
```

## Cluster 26
```{r}
clusterVis(26)
clusterVis_color(26)
clusterVis_PCA(26) 
```

## Cluster 27
```{r}
clusterVis(27)
clusterVis_color(27)
clusterVis_PCA(27) 
```

## Cluster 28
```{r}
clusterVis(28)
clusterVis_color(28)
clusterVis_PCA(28) 
```

## Cluster 29
```{r}
clusterVis(29)
clusterVis_color(29)
clusterVis_PCA(29) 
```

## Cluster 30
```{r}
clusterVis(30)
clusterVis_color(30)
clusterVis_PCA(30) 
```

## Cluster 31
```{r}
clusterVis(31)
clusterVis_color(31)
clusterVis_PCA(31) 
```

## Cluster 32
```{r}
clusterVis(32)
clusterVis_color(32)
clusterVis_PCA(32) 
```

## Cluster 33
```{r}
clusterVis(33)
clusterVis_color(33)
clusterVis_PCA(33) 
```

## Cluster 34
```{r}
clusterVis(34)
clusterVis_color(34)
clusterVis_PCA(34) 
```

## Cluster 35
```{r}
clusterVis(35)
clusterVis_color(35)
clusterVis_PCA(35) 
```

## Cluster 36
```{r}
clusterVis(36)
clusterVis_color(36)
clusterVis_PCA(35) 
```

### Making dataframe for output
```{r}
annotation2 <- read.delim("../../../06diffGeneExp/analysis4_1Sept2017/data/ITAG2.3_all_Arabidopsis_annotated.tsv")

annotation2$ITAG <- gsub("^(.*)[.].*", "\\1", annotation2$ITAG)
annotation2$ITAG <- gsub("^(.*)[.].*", "\\1", annotation2$ITAG)

annotation2 <- annotation2[,c(1:4)]

###Output for DE promoter enrichment.
data.valOutput <- data.val2
names(data.valOutput)[20:21] <- c("cluster", "distanceA")
names(data.valOutput)[1] <- "ITAG"

# subset only genes and cluster number
data.valOutput <- data.valOutput[,c(1, 20)]
## write.csv(data.valOutput, "../data/output/analysis11.large.03282018_with_annotations_tf2.csv", row.names = FALSE)

## Merge

head(annotation2)
dim(annotation2) #14640
head(data.valOutput)
dim(data.valOutput) #6582

data.valOutput_2 <- merge(data.valOutput, annotation2, by = "ITAG", all.x = TRUE )

dim(data.valOutput_2) #6582

# write.csv(data.valOutput_2, "../data/output/analysis11.large.100917_with_annotations.csv_tf2", row.names = FALSE)

data.valOutput_2 <- read.csv("../data/output/analysis11.large.100917_with_annotations.csv")
```
